---
title: "PCAngsd_Analysis"
author: "Alyssa Green"
date: "`r Sys.Date()`"
output: html_document
---
Code based from: PCAngsdTutorial—Software. (n.d.). Retrieved February 4, 2023, from http://www.popgen.dk/software/index.php/PCAngsdTutorial

Load Libraries
```{r}
library(RcppCNPy)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(grid)
library(cowplot)
source("http://bit.ly/theme_pub") # By Dr. Rob Colautti
```

Load Data
```{r}
#Load in .npy file
npy_dat<- npyLoad("try2_selection.selection.npy")

#Load in .sites file
sites<-read.table("kittiwake.sites", row.names = NULL)
names(sites)<-c("chromosome","marker")

#Load in alternate 0/1 .sites file
alt_sites<-read.table("try2_selection.sites")

#Load Genes File
Genes=read.csv("Genes.csv")
```

Create QQplot to QC Test Statistics
```{r}
#Run qqchi function
source("./qqchi_func.R")
qqchi(npy_dat)
```

Convert test statistics to p-values
```{r}
pval=1-pchisq(npy_dat,1)
pval=as.data.frame(pval)
```

Combine 2 .sites files
```{r}
sites$V1=alt_sites$V1
```

Remove elements of .sites values where V1=0
```{r}
sig=sites%>%filter(V1==1)
```

Create data frame combining sites and pvals
```{r}
Combo=data.frame(marker=sig$marker, chromosome=sig$chromosome, pval1=pval$V1, pval2=pval$V2)
```

Select markers of interest
```{r}
to_keep = c()
gene_annotation = c()
scaffold=c()
p1=c()
p2=c()
group=c()


  for (i in 1:length(Combo$chromosome)) {
  for (j in 1:length(Genes$Code)){
      if (Genes$Code[j] == Combo$chromosome[i]){
        if(between(Combo$marker[i], Genes$Min[j], Genes$Max[j])){
            to_keep = c(to_keep,i)
            gene_annotation = c(gene_annotation,Genes$Gene[j])
            scaffold=c(scaffold, Combo$chromosome[i])
            p1=c(p1, Combo$pval1[i])
            p2=c(p2, Combo$pval2[i])
            group=c(group, Genes$Group[j])
            
              }
         }
    }
}

```

Create data frame with information for genes on interest
```{r}
Data=data.frame(gene_annotation, scaffold, marker=to_keep, p1, p2, group)
head(Data)

```

Manhattan plots

P1 all
```{r, fig.width=10, fig.height=2}
ggplot(aes(x=gene_annotation, y=-log10(p1), colour=group), data=Data)+geom_point()+theme_pub()+geom_hline(yintercept=5.4)+xlab("Gene")+ylab("-log10(p-value)")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=4))+ theme(legend.position = "right")+scale_color_discrete(labels=c("Cytochrome B", "DNA Polymerase", "GRE", "RNA Polymerase","Telomerase", "Thyroid"))+guides(color=guide_legend(title="Gene Group"))

pp1=ggplot(aes(x=gene_annotation, y=-log10(p1), colour=group), data=Data)+geom_point()+theme_pub()+geom_hline(yintercept=5.4, colour="hotpink")+theme(axis.title.x = element_blank(), axis.title.y = element_blank())+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))+theme(plot.title = element_text(hjust = 0))+ggtitle("A")
```


P2 all
```{r, fig.width=10, fig.height=2}
ggplot(aes(x=gene_annotation, y=-log10(p2), colour=group), data=Data)+geom_point()+theme_pub()+geom_hline(yintercept=5.4)+xlab("Gene")+ylab("-log10(p-value)")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=4))+ theme(legend.position = "right")+scale_color_discrete(labels=c("Cytochrome B", "DNA Polymerase", "Glucocorticoid", "RNA Polymerase","Telomerase", "Thyroid"))+guides(color=guide_legend(title="Gene Group"))

pp2=ggplot(aes(x=gene_annotation, y=-log10(p2), colour=group), data=Data)+geom_point()+theme_pub()+geom_hline(yintercept=5.4, colour="hotpink")+theme(axis.title.x = element_blank(), axis.title.y = element_blank())+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))+theme(plot.title = element_text(hjust = 0))+ggtitle("B")
```


```{r, fig.width=10, fig.height=4}
grid.arrange(pp1, pp2, ncol=1, bottom=textGrob("Gene", gp=gpar(fontsize=20)), left=textGrob("-log10(p-value)", gp=gpar( fontsize=20), rot=90))
```

